---
title: "Analysis by funtional group per depth within site"
author: "Claudia Campanini"
date: '2022-07-26'
output: 
  html_document: 
      toc: yes
      number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r libraries}
library(readxl) #read xlsx files
library(tidyverse) #aggregate fc et al.
library(ggplot2)
library(dplyr)
library(dplyr)
library(MASS)
library(here)
```

# Stacked bar chart

Load the input dataframe

```{r}
func_mean_depth<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_mean_depth_input")

str(func_mean_depth) 
func_mean_depth$Site <- ordered(func_mean_depth$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))
func_mean_depth$Depth[func_mean_depth$Depth == "Deep"] <- "D"
func_mean_depth$Depth[func_mean_depth$Depth == "Shallow"] <- "S"
func_mean_depth[,c(2:3)] <- lapply(func_mean_depth[,c(2:3)], as.factor) 

str(func_mean_depth)

func_mean_depth$Functional_group[func_mean_depth$Functional_group == "plate"] <- "solitary"
```
Add the pressure column

```{r}
library(magrittr)
func_mean_depth %<>%
  mutate(Pressure = case_when(Site == 'HI' | 
                                Site == "S1" | 
                                Site == "S2" ~ "Low",
                              Site == "ZP" |
                                Site == 'ZA' | 
                                Site == "YO" | 
                                Site == "MI" ~ "Medium",
                              Site == 'SK' |
                                Site == 'SN' | 
                                Site == "AW" | 
                                Site == "G1" |
                                Site == "G2"~ "High"))
```


## Hard coral stacked bar chart

```{r}
hc_mean_depth <- subset(func_mean_depth,func_mean_depth$Category == "hard_coral")

hc_mean_depth$Functional_group <- ordered(hc_mean_depth$Functional_group, levels= c("encrusting", "solitary", "submassive", "massive", "laminar","table", "foliose", "digitate","corymbose", "hispidose" , "branching"))
```



```{r}
#remotes::install_github("wilkelab/cowplot")
#install.packages("colorspace")
library(colorspace)

#install.packages("colorspace")
#remotes::install_github("clauswilke/colorblindr")
library(colorblindr)

palette10plasma=c("#f0f921","#fdca26","#fb9f3a","#ed7953","#d8576b",
                  "#bd3786","#9c179e","#7201a8","#46039f","#0d0887")

plasma5 = sequential_hcl(5, palette= "plasma")
plasma5
ag_GrnYl5 = sequential_hcl(5, palette= "ag_GrnYl")
ag_GrnYl5
sequential_hcl(1, palette= "Emrld") # --> "#0B4151"

  
plasma_ag_GrnYl_11 = c("#46039f","#9c179e","#d8576b","#ed7953","#EDA200","gold1", #plasma
                        "#EDEF5C","#82CC6C","#17A77E","#007E7D","#0B4151")

ggplot(hc_mean_depth, 
       aes(x=Depth, y=mean_cat_tot/100, fill=Functional_group)) +
  geom_bar(position="stack",stat="identity", color = "black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [m]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_fill_manual(values=plasma_ag_GrnYl_11)+
  scale_y_continuous(breaks = seq(0,15,3))
```

```{r}
palette11plasma=c("#f0f921","gold1","#fdca26","#fb9f3a","#ed7953","#d8576b",
                  "#bd3786","#9c179e","#7201a8","#46039f","#0d0887")

ggplot(hc_mean_depth, 
       aes(x=Depth, y=mean_cat_tot/100, fill=Functional_group)) +
  geom_bar(position="stack",stat="identity", color = "black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [m]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_fill_manual(values=palette11plasma)+
  scale_y_continuous(breaks = seq(0,15,3))
```
```{r}
plasma_ag_GrnYl_11 = c("#001889","#91008D","#D24E71","#EDA200","gold1", "#EDEF5C","#82CC6C","#17A77E","#007E7D","#255668","#0B4151")

hc_mean_depth$Functional_group <- as.character(hc_mean_depth$Functional_group)

hc_mean_depth$Functional_group <- as.factor(hc_mean_depth$Functional_group)

hc_mean_depth$Functional_group <- ordered(hc_mean_depth$Functional_group, levels= c("encrusting", "solitary", "submassive", "massive", "laminar","table", "foliose", "digitate","corymbose", "hispidose" , "branching"))

ggplot(hc_mean_depth, 
       aes(x=Depth, y=mean_cat_tot/100, fill=Functional_group)) +
  geom_bar(position="stack",stat="identity", color = "black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [m]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_fill_manual(values=plasma_ag_GrnYl_11)+
  scale_y_continuous(breaks = seq(0,15,3))
```



## Hard coral stacked bar chart in percentage

See palette from
<https://cran.r-project.org/web/packages/colorspace/vignettes/colorspace.html#Usage_with_ggplot2>

```{r}
# install.packages("ggh4x")
library(ggh4x) # needed for nested faceting

hc_func_stacked <- ggplot(subset(func_mean_depth,func_mean_depth$Category == "hard_coral"), 
       aes(x=Depth, y=mean_cat_tot, fill=Functional_group)) +
  geom_bar(position="fill",stat="identity",color="black")+
  facet_nested(~ Pressure + Site, scales = "free_x", space = "free_x") +
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [%]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_fill_manual(values=plasma_ag_GrnYl_11)+
  scale_y_continuous(label = c("0","25","50","75","100"))

hc_func_stacked
```


```{r, eval = F}
ggsave("hard_coral_growth_forms_stacked.tiff", 
       plot = hc_func_stacked,
       units = "in", width = 7, height = 4.5, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))

ggsave("hard_coral_growth_forms_stacked.png", 
       plot = hc_func_stacked,
       units = "in", width = 7, height = 4.5, dpi = 300,
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# Algae stacked bar chart

```{r}
alg_mean_depth <- subset(func_mean_depth,func_mean_depth$Category == "algae")

alg_mean_depth <- alg_mean_depth %>% 
  group_by(Depth,Site) %>% 
  mutate(Perc_func = round(mean_cat_tot/sum(mean_cat_tot)*100, 0))

RedGreen5 = colorspace::diverging_hcl(5, palette= "Red-Green")
RedGreen3 = c("#841859","#7CC57D","#005600")

algae_func_stacked <- ggplot(alg_mean_depth,aes(x=Depth, y=mean_cat_tot, fill=Functional_group)) +
  geom_bar(position="fill",stat="identity",alpha=0.7,color="black")+
  geom_text(aes(x=Depth, y= mean_cat_tot, 
                label = Perc_func), 
            data = subset(alg_mean_depth,alg_mean_depth$Perc_func > 2),
            size = 2.5, 
            position = position_fill(vjust = 0.5),color="black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean algae cover by functional group [%]")+
  guides(fill=guide_legend(title="Functional groups"))+
  scale_fill_manual(values=RedGreen3)+
  scale_y_continuous(label=c(0,25,50,75,100))

cvd_grid(algae_func_stacked)

algae_func_stacked
```
```{r, eval = F}
ggsave("algae_func_stacked.tiff", 
       units = "in", width = 7, height = 4.5, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```



```{r}
algae_func_bar <- 
  ggplot(alg_mean_depth,
         aes(x=Depth, y=mean_cat_tot/100, fill=Functional_group)) +
  geom_bar(position="stack",stat="identity",alpha=0.7,color="black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean algae cover by functional group [m]")+
  guides(fill=guide_legend(title="Functional groups"))+
  scale_y_continuous(breaks = seq(0,25,5),limits=c(0,25))+
  scale_fill_manual(values=RedGreen3)

algae_func_bar
```

```{r, eval = F}
ggsave("algae_func_bar.tiff", 
       units = "in", width = 7, height = 4.5, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

```{r, eval = F}
ggsave("algae_func_bar.png", 
       units = "in", width = 7, height = 4.5, dpi = 300,
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# nMDS

Load packages

```{r}
library(here)
library(vegan)#load package
library(MASS)
library(dplyr)
library(readxl)   # to import excel datasets
```

Load input dataframe

```{r}
func_input<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_input")

func_input <- func_input %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

func_input <- func_input[ ,c(1:5, 20, 6:19)]

func_input[,c(1:6)] <- lapply(func_input[,c(1:6)], as.factor)

func_input$Pressure<-ordered(func_input$Pressure,levels=c("Low", "Medium", "High"))

func_input$Site <- ordered(func_input$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

str(func_input)


#create a df with the independent variables
ind_df <- as.data.frame(func_input[,c(1:6)])
str(ind_df)

```

Convert the database `func_input`to a matrix

```{r}
func_mat <- func_input[,-c(1:6)]
rownames(func_mat) <- func_input$Trans
func_mat <- as.matrix(func_mat)
str(func_mat)
class(func_mat)
```

## Run the Nonmetric Multidimensional scaling

```{r}
##metaMDS (isoMDS with random start, reduced  stress)----
set.seed(69) #random number to start
func_nmds_br=metaMDS(func_mat,distance = "bray",k=2) #Bray-Curtis resemblance matrix;k=2 means 2 dimensional
```

## Evaluate the nMDS mapping based on the stress and goodness of fit

```{r}
stressplot(func_nmds_br) #evaluating NMDS mapping: stressplot
func_nmds_br_gof=goodness(func_nmds_br) #evaluating NMDS mapping: goodness of fit
func_nmds_br_gof #smaller the value, the lower stress and better fit
plot(func_nmds_br, type="t", main="goodness of fit") # create a plot for goodness of fit visualization:
points(func_nmds_br, display="site", cex=func_nmds_br_gof*100) #display the goodness of fit on the plot
#Stress lesser, better (>0.2 poor,0.2-0.1 moderate,0.1-0.05 good,<0.05 excellent); 
#K level higher, stress lesser
```

Stress = `r func_nmds_br$stress` \< 0.2

Check Stress for each additional dimension

```{r}
k_vec=1:10
stress=numeric(length(k_vec))
relfunc_dij=metaMDSdist(func_mat, trace=FALSE)
set.seed(69)
for(i in seq_along(k_vec)) {
  func_nmds_br_loop=metaMDSiter(relfunc_dij, k=i, trace=FALSE)
  stress[i]=func_nmds_br_loop$stress
}
plot(k_vec,stress,type="b",ylab="Stress",xlab="Dimensions") #select the K correspondent to the most drastic drop in stress
```

## nMDS plot with `ggplot()` and `geom_mark_ellipse()` from `ggforce`

N.B. Visualisation with ggplot()\` first require the creation of an
input dataframe with nMDS scores

```{r}
scores(func_nmds_br)
func_ndms_scores <- as.data.frame(scores(func_nmds_br)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
func_ndms_scores$transect <- rownames(func_ndms_scores) # create a column of transect names
func_ndms_scores$site <- ind_df$site
func_ndms_scores$depth <- ind_df$depth  #  add the grp variable created earlier
str(func_ndms_scores)  #look at the data


NMDS1 <- func_nmds_br$points[,1] ##also found using: > scores(faMDS)
NMDS2 <- func_nmds_br$points[,2]
nmds_df<-cbind(ind_df, NMDS1, NMDS2)

```

nMDS plot with colorblind-friendly

```{r}
#install.packages("ggforce")
library(ggforce)
library(ggplot2)
```

Use

```{r}
pal12 <- c("#000000","#004949","#009292","#ffb6db",
 "#490092","#006ddb","#b66dff","#b6dbff",
 "#920000","#924900","#24ff24","#ffff6d")

library(ggnewscale) #necessary for new_scale_color 
```


```{r}
pal_fabian <- c("#8400CD","#00C2F9","#FFB2FD","#009F81","#FFC33B","#FF6E3A","#008DF9","#A40122","#E20134","#9F0162","#FF5AAF","#00FCCF")

func_ndms_ggplot <- ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Pressure),size=2.5)+##separates overlapping points
  scale_color_manual(values= pal_fabian)+
  guides(color= guide_legend(order=1, title = "Site", ncol =2))+
  new_scale_color() +
  geom_mark_ellipse(aes(fill=Pressure, color=Pressure), expand = 0, alpha=0.2)+ # https://luisdva.github.io/rstats/Grouping-points/
  scale_color_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  scale_fill_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  theme_classic()+
  ggtitle("nMDS of relative cover of hard coral and algae \n by functional group")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  scale_x_continuous(limits = c(-1.5,0.8))+
  scale_y_continuous(limits = c(-0.5,.6))+
  annotate("text", x=-1.45, y=-0.45, 
           label=bquote(atop(Stress == .(round(func_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.921")),
           size=3, hjust=0) #add goodness of fit to plot


func_ndms_ggplot

rm(pal12)
```

```{r}
func_ndms_ggplot<- ggplot(nmds_df, aes(NMDS1, NMDS2))+
  stat_ellipse(geom = "polygon", aes(fill = Pressure), alpha = 0.2) +
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Depth),size=2.5, alpha = 0.8)+##separates overlapping points+
  scale_fill_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  scale_color_manual(values= pal_fabian)+
  labs(x = "NMDS1", colour = "Site", y = "NMDS2") +
  guides(color= guide_legend(order=1, title = "Site", ncol =2),
         fill=guide_legend(order = 2))+
  theme_classic()+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
    annotate("text", x=Inf, y=-Inf, hjust = 1, vjust = -0.5,
           label=bquote(atop(Stress == .(round(func_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.993")),
           size=3) #add goodness of fit to plot

func_ndms_ggplot
```

Save the plot 

```{r, eval = F}
ggsave("func_nmds.tiff", 
       plot = func_ndms_ggplot,
       units="in", width=6, height=5, dpi=600, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs")
       )

ggsave("func_nmds.png", 
       plot = func_ndms_ggplot,
       units="in", width=6, height=5, dpi=600,
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs")
       )
```

## Figure 3

```{r}
fig3 <- cowplot::  ggdraw() +
  draw_plot(hc_func_stacked, x = 0, y = 0, width = .6, height = 1) +
  draw_plot(func_ndms_ggplot, x = .6, y = 0, width = 0.4, height = 1) +
  draw_plot_label(label = c("a)", "b)"), size = 14,
                  x = c(0, 0.6), y = c(1, 1))

fig3
```

Save the plot 

```{r, eval = F}
ggsave("fig3.tiff", 
       plot = fig3,
       units ="in", width = 11, height = 5, dpi = 600, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))

ggsave("fig3.png", 
       plot = fig3,
       units ="in", width = 11, height = 5, dpi = 600,
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# MANTEL TEST

```{r load libraries for Mantel test}
library(fields)
library(reshape2)
library(ade4)
library(adespatial)
library(spdep)
library(ggplot2)
```

Load distance file and create distance matrix 
```{r load distance file}
Ok_dist_69 <- read.csv(file=here("Multi-taxa_data","Ok_dist_69.csv"), row.names=1 )
head(Ok_dist_69)
Dist_km <- as.dist(rdist.earth(Ok_dist_69, miles=F))

func_bray <- vegdist(func_mat, method = "bray")
```


Correlation between straight-line geographic distance and dissimilarity in species composition

```{r}
set.seed(69)
mantel(Dist_km, func_bray)
```

Plot
```{r}
func_matrix_dist <- data.frame(x = melt(as.matrix(Dist_km))$value, 
                          y = melt(as.matrix(func_bray))$value) ## Prepare data for graphic

ggplot(func_matrix_dist) +
  geom_jitter( aes(x,y), shape = 21, size = 3, fill = "#8fc7a0") +
  labs(x = "Distance (km)", 
       y = "Dissimilarity (Bray-Curtis)") +
  geom_smooth( aes(x,y), method = "glm") +
  theme_classic()
```
```{r, eval=F}
ggsave("func_dissimil_distance.tiff", 
       units="in", width=6, height=4, dpi=300, compression = 'lzw',
       path = here ("Multi-taxa_data","PLITs","4-Visualization_outputs"))

ggsave("func_dissimil_distance.png", 
       units="in", width=6, height=4, dpi=300,
       path = here ("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

Mantel 
```{r}
set.seed(69)
mantel_func <- mantel.randtest(func_bray, Dist_km)
mantel_func
```

# Boxplot with within-site distance from centroids made with `ggplot()`

Load packages

```{r}
library(here)
library(vegan)#load package
library(MASS)
library(dplyr)
library(readxl)   # to import excel datasets
```

Load input dataframe

```{r}
func_input<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_input")

func_input <- func_input %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

func_input <- func_input[ ,c(1:5, 20, 6:19)]

func_input[,c(1:6)] <- lapply(func_input[,c(1:6)], as.factor)

func_input$Pressure<-ordered(func_input$Pressure,levels=c("Low", "Medium", "High"))

func_input$Site <- ordered(func_input$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

str(func_input)


#create a df with the independent variables
ind_df <- as.data.frame(func_input[,c(1:6)])
str(ind_df)

```
Convert the database `func_input`to a matrix

```{r}
func_mat <- func_input[,-c(1:6)]
rownames(func_mat) <- func_input$Trans
func_mat <- as.matrix(func_mat)
str(func_mat)
class(func_mat)
```

Create an input dataframe for the boxplot with mean relative cover by
category and depth within site

```{r}
func_dist = vegdist(func_mat, 
                   method="bray")
set.seed(69)
func_bd_site = betadisper(func_dist, 
                    ind_df$Site)  #Implements Marti Anderson's PERMDISP2 procedure for the analysis of multivariate homogeneity of group dispersions (variances). betadisper is a multivariate analogue of Levene's test for homogeneity of variances 

str(func_bd_site)

func_bd_site$distances #Extract betadisper scores

distcent_df <- data.frame(Distance_to_centroid=func_bd_site$distances,
                          Site=func_bd_site$group) #create a df for the plot


distcent_df <- distcent_df %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

distcent_df[,2]<-ordered(distcent_df[,2],levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

distcent_df$Pressure<-ordered(distcent_df$Pressure,levels=c("Low", "Medium", "High"))

distcent_df_mean <- distcent_df %>%
  group_by(Site, Pressure) %>%
  summarise(Mean_dist = round(mean(Distance_to_centroid),2),
            SD = round(sd(Distance_to_centroid),2)) %>%
  ungroup()


colnames(distcent_df_mean) <- c("Site","Pressure","Mean distance to centroid","SD")
```

Boxplot with `geom_boxplot` from `ggplot2`

```{r}
box<- ggplot(data=distcent_df,aes(x=Site,y=Distance_to_centroid))+ 
  geom_boxplot(aes(fill=Pressure,colour=Pressure),alpha=0.3)+   
  geom_jitter(aes(color=Pressure),alpha=0.7)+
  theme_classic()+
  ylab(label="Distance to centroid")+
  theme(legend.position = "none")+ #remove legend
  scale_fill_manual(values=c("#009F81", "#FFB331","#9F1111"))+
  scale_color_manual(values=c("#009F81", "#FFB331","#9F1111"))
box  

ggpubr:: ggarrange(box, 
                   gridExtra::tableGrob(distcent_df_mean), 
                   ncol = 2, widths = c(1,1))
```


```{r, eval = F}
ggsave("func_boxplot.tiff",
       units="in", width=12, height=4, dpi=300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))

ggsave("func_boxplot.png",
       units="in", width=12, height=4, dpi=300,
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# SIMPER: Similarity of percentage

From the vignette: *The simper functions performs pairwise comparisons
of groups of sampling units and finds the contribution of each species
to the average between-group Bray-Curtis dissimilarity. Although the
method is called simper, it really studied dissimilarities instead of
similarities (Clarke 1993).*

## SIMPER PER SITE:

```{r}
func_sim_site=simper(func_mat,group=ind_df$Site) #Distinguish categories which caused the differences in relative cover among sites
func_sim_site
summary(func_sim_site) #check which categories were the main driver of difference among sites
```

## SIMPER per depth

```{r}
func_sim_depth=simper(func_mat,group=ind_df$Depth) #Distinguish categories which caused the differences in relative cover among sites
func_sim_depth
summary(func_sim_depth, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

## SIMPER per pressure exposure
```{r}
func_sim_pres=simper(func_mat,group=ind_df$Pressure) #Distinguish categories which caused the differences in relative cover among sites
func_sim_pres
summary(func_sim_pres, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

# Session info

```{r}
sessionInfo()
```

# References

```{r}
#install.packages("report")
library(report)
report(sessionInfo())
```
