---
title: "Analysis by funtional group per depth within site"
author: "Claudia Campanini"
date: '2022-07-26'
output: 
  html_document: 
      toc: yes
      number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r libraries}
library(readxl) #read xlsx files
library(tidyverse) #aggregate fc et al.
library(ggplot2)
library(dplyr)
library(dplyr)
library(MASS)
library(here)
```

# Stacked bar chart

Load the input dataframe

```{r}
func_mean_depth<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_mean_depth_input")

str(func_mean_depth) 
func_mean_depth$Site <- ordered(func_mean_depth$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))
func_mean_depth$Depth[func_mean_depth$Depth == "Deep"] <- "D"
func_mean_depth$Depth[func_mean_depth$Depth == "Shallow"] <- "S"
func_mean_depth[,c(2:4)] <- lapply(func_mean_depth[,c(2:4)], as.factor) 

str(func_mean_depth)

```

## Hard coral stacked bar chart

```{r}

#remotes::install_github("wilkelab/cowplot")
#install.packages("colorspace")
library(colorspace)

#install.packages("colorspace")
#remotes::install_github("clauswilke/colorblindr")
library(colorblindr)

palette10plasma=c("#f0f921","#fdca26","#fb9f3a","#ed7953","#d8576b",
                  "#bd3786","#9c179e","#7201a8","#46039f","#0d0887")

plasma5 = sequential_hcl(5, palette= "plasma")
plasma5
ag_GrnYl5 = sequential_hcl(5, palette= "ag_GrnYl")
ag_GrnYl5
sequential_hcl(1, palette= "Emrld") # --> "#0B4151"

  
plasma_ag_GrnYl_11 = c("#001889","#91008D","#D24E71","#EDA200","gold1", #plasma
                        "#EDEF5C","#82CC6C","#17A77E","#007E7D","#255668","#0B4151")

ggplot(subset(func_mean_depth,func_mean_depth$Category == "hard_coral"), 
       aes(x=Depth, y=mean_cat_tot, fill=Functional_group)) +
  geom_bar(position="stack",stat="identity", color = "black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [m]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_y_continuous(label = c("0","5","10","15"))+
  scale_fill_manual(values=plasma_ag_GrnYl_11)
```

## Hard coral stacked bar chart in percentage

See palette from
<https://cran.r-project.org/web/packages/colorspace/vignettes/colorspace.html#Usage_with_ggplot2>

```{r}
hc_func_stacked <- ggplot(subset(func_mean_depth,func_mean_depth$Category == "hard_coral"), 
       aes(x=Depth, y=mean_cat_tot, fill=Functional_group)) +
  geom_bar(position="fill",stat="identity",color="black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean hard coral cover by growth form [%]")+
  guides(fill=guide_legend(title="Growth form"))+
  scale_fill_manual(values=plasma_ag_GrnYl_11)+
  scale_y_continuous(label = c("0","25","50","75","100"))

cvd_grid(hc_func_stacked)

hc_func_stacked
```
```{r, eval = F}
ggsave("hard_coral_growth_forms_stacked.tiff", 
       units = "in", width = 7, height = 4.5, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# Algae stacked bar chart

```{r}
alg_mean_depth <- subset(func_mean_depth,func_mean_depth$Category == "algae")

alg_mean_depth <- alg_mean_depth %>% 
  group_by(Depth,Site) %>% 
  mutate(Perc_func = round(mean_cat_tot/sum(mean_cat_tot)*100, 0))

RedGreen5 = diverging_hcl(5, palette= "Red-Green")
RedGreen3 = c("#841859","#7CC57D","#005600")

algae_func_stacked <- ggplot(alg_mean_depth,aes(x=Depth, y=mean_cat_tot, fill=Functional_group)) +
  geom_bar(position="fill",stat="identity",alpha=0.7,color="black")+
  geom_text(aes(x=Depth, y= mean_cat_tot, label = Perc_func), 
            data = subset(alg_mean_depth,alg_mean_depth$Perc_func > 2),
            size = 2.5, position = position_fill(vjust = 0.5),color="black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean algae cover by functional group [%]")+
  guides(fill=guide_legend(title="Functional groups"))+
  scale_fill_manual(values=RedGreen3)+
  scale_y_continuous(label=c(0,25,50,75,100))

cvd_grid(algae_func_stacked)

algae_func_stacked
```


```{r, eval = F}
ggsave("algae_func_stacked.tiff", 
       units = "in", width = 7, height = 4.5, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# nMDS

Load packages

```{r}
library(vegan)#load package
library(MASS)
library(dplyr)
library(readxl)   # to import excel datasets
```

Load input dataframe

```{r}
func_input<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_input")

func_input <- func_input %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

func_input <- func_input[ ,c(1:5, 20, 6:19)]

func_input[,c(1:6)] <- lapply(func_input[,c(1:6)], as.factor)

func_input$Pressure<-ordered(func_input$Pressure,levels=c("Low", "Medium", "High"))

func_input$Site <- ordered(func_input$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

str(func_input)


#create a df with the independent variables
ind_df <- as.data.frame(func_input[,c(1:6)])
str(ind_df)

```

Convert the database `func_input`to a matrix

```{r}
func_mat <- func_input[,-c(1:6)]
rownames(func_mat) <- func_input$Trans
func_mat <- as.matrix(func_mat)
str(func_mat)
class(func_mat)
```

## Run the Nonmetric Multidimensional scaling

```{r}
##metaMDS (isoMDS with random start, reduced  stress)----
set.seed(69) #random number to start
func_nmds_br=metaMDS(func_mat,distance = "bray",k=2) #Bray-Curtis resemblance matrix;k=2 means 2 dimensional
```

## Evaluate the nMDS mapping based on the stress and goodness of fit

```{r}
stressplot(func_nmds_br) #evaluating NMDS mapping: stressplot
func_nmds_br_gof=goodness(func_nmds_br) #evaluating NMDS mapping: goodness of fit
func_nmds_br_gof #smaller the value, the lower stress and better fit
plot(func_nmds_br, type="t", main="goodness of fit") # create a plot for goodness of fit visualization:
points(func_nmds_br, display="site", cex=func_nmds_br_gof*100) #display the goodness of fit on the plot
#Stress lesser, better (>0.2 poor,0.2-0.1 moderate,0.1-0.05 good,<0.05 excellent); 
#K level higher, stress lesser
```

Stress = `r func_nmds_br$stress` \< 0.2

Check Stress for each additional dimension

```{r}
k_vec=1:10
stress=numeric(length(k_vec))
relfunc_dij=metaMDSdist(func_mat, trace=FALSE)
set.seed(69)
for(i in seq_along(k_vec)) {
  func_nmds_br_loop=metaMDSiter(relfunc_dij, k=i, trace=FALSE)
  stress[i]=func_nmds_br_loop$stress
}
plot(k_vec,stress,type="b",ylab="Stress",xlab="Dimensions") #select the K correspondent to the most drastic drop in stress
```

## nMDS plot with `ggplot()` and `geom_mark_ellipse()` from `ggforce`

N.B. Visualisation with ggplot()\` first require the creation of an
input dataframe with nMDS scores

```{r}
scores(func_nmds_br)
func_ndms_scores <- as.data.frame(scores(func_nmds_br)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
func_ndms_scores$transect <- rownames(func_ndms_scores) # create a column of transect names
func_ndms_scores$site <- ind_df$site
func_ndms_scores$depth <- ind_df$depth  #  add the grp variable created earlier
str(func_ndms_scores)  #look at the data


NMDS1 <- func_nmds_br$points[,1] ##also found using: > scores(faMDS)
NMDS2 <- func_nmds_br$points[,2]
nmds_df<-cbind(ind_df, NMDS1, NMDS2)

```

nMDS plot with colorblind-friendly

```{r}
#install.packages("ggforce")
library(ggforce)
library(ggplot2)
```

Use

```{r}
pal12 <- c("#000000","#004949","#009292","#ffb6db",
 "#490092","#006ddb","#b66dff","#b6dbff",
 "#920000","#924900","#24ff24","#ffff6d")

library(ggnewscale) #necessary for new_scale_color 

ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Depth),size=2.5)+##separates overlapping points
  scale_color_manual(values= pal12)+
  guides(color= guide_legend(order=1, title = "Site", ncol =2))+
  new_scale_color() +
  geom_mark_ellipse(aes(fill=Pressure, color=Pressure), expand = 0, alpha=0.2)+ # https://luisdva.github.io/rstats/Grouping-points/
  scale_color_manual(values= c("green","yellow","red"))+
  scale_fill_manual(values= c("green","yellow","red"))+
  theme_classic()+
  ggtitle("nMDS of relative cover of hard coral and algae \n by functional group")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  scale_x_continuous(limits = c(-1.5,0.8))+
  scale_y_continuous(limits = c(-0.5,.6))+
  annotate("text", x=-1.4, y=-0.45, 
           label=bquote(atop(Stress == .(round(func_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.921")),
           size=3, hjust=0) #add goodness of fit to plot

```

```{r}
func_ndms_ggplot <- ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Pressure),size=2.5)+##separates overlapping points
  scale_color_manual(values= pal12)+
  guides(color= guide_legend(order=1, title = "Site", ncol =2))+
  new_scale_color() +
  geom_mark_ellipse(aes(fill=Pressure, color=Pressure), expand = 0, alpha=0.2)+ # https://luisdva.github.io/rstats/Grouping-points/
  scale_color_manual(values= c("green","yellow","red"))+
  scale_fill_manual(values= c("green","yellow","red"))+
  theme_classic()+
  ggtitle("nMDS of relative cover of hard coral and algae \n by functional group")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  scale_x_continuous(limits = c(-1.5,0.8))+
  scale_y_continuous(limits = c(-0.5,.6))+
  annotate("text", x=-1.45, y=-0.45, 
           label=bquote(atop(Stress == .(round(func_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.921")),
           size=3, hjust=0) #add goodness of fit to plot


func_ndms_ggplot

rm(pal12)
```


Test a colorblind-friendly palette using `colorblindr` package

```{r}


#remotes::install_github("wilkelab/cowplot")
#install.packages("colorspace")
library(colorspace)

#install.packages("colorspace")
#remotes::install_github("clauswilke/colorblindr")
library(colorblindr)

cvd_grid(func_ndms_ggplot)

```

Save the plot as tiff

```{r, eval = F}
ggsave("func_nmds.tiff", 
       plot = func_ndms_ggplot,
       units="in", width=7, height=4.5, dpi=300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs")
       )
```

# PERMANOVA

Load packages

```{r}
#VEGAN----
library(vegan)#load package
#install.packages("vegan3d")
library(MASS)
library(dplyr)

#install.packages("pairwiseAdonis")
library(devtools) #to install pairwiseAdonis from GitHub
library(pairwiseAdonis)
library(readxl)   # to import excel datasets
```

Load input data frame

```{r}
func_input<- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","func_clean.xlsx"), sheet = "func_input")

func_input <- func_input %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

func_input <- func_input[ ,c(1:5, 20, 6:19)]

func_input[,c(1:6)] <- lapply(func_input[,c(1:6)], as.factor)

func_input$Pressure<-ordered(func_input$Pressure,levels=c("Low", "Medium", "High"))

func_input$Site <- ordered(func_input$Site, levels= c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

str(func_input)


#create a df with the independent variables
ind_df <- as.data.frame(func_input[,c(1:6)])
str(ind_df)


```

Convert the database `func_input`to a matrix

```{r}
func_mat <- func_input[,-c(1:6)]
rownames(func_mat) <- func_input$Trans
func_mat <- as.matrix(func_mat)
str(func_mat)
class(func_mat)
```

## PERMANOVA among the 4 sites

### Question 1: is there a significant difference in relative algae and coral cover by functional groups among sites?

```{r}
set.seed(69) 
perm=how(nperm=999)
func_site_pmv=adonis2(func_mat~Site, 
                        data=ind_df,
                        permutations =perm,
                        method="bray",
                      by = "margin")
                        
func_site_pmv
```

P-value `r func_site_pmv$Pr[1]` \< 0.05, so there is a significant
difference among sites.

### Question 2: is there a significant difference in relative algae and coral cover by functional group between depth withing sites?

```{r}
#SCRIPT WITH adonis()
set.seed(69) 
perm=how(nperm=999)
func_depth_pmv <- adonis(func_mat~Depth, 
                           data=ind_df,
                           permutations =perm,
                           strata=ind_df$Site,
                           method="bray",
                           by = "margin")
func_depth_pmv$aov.tab

#ALTERNATIVE SCRIPT WITH adonis2()
set.seed(69) 
h <- how(blocks = ind_df$Site, nperm = 999)
func_blocks_pmv <- adonis2(func_mat~Depth,
                             data=ind_df,
                             permutations =h,
                             method="bray",
                             by = "margin")

func_blocks_pmv
```

### Question 3: is there a significant difference in relative cover by functional group among anthropogenic pressure level?

```{r}
set.seed(69)
perm=how(nperm=999)
func_pressure_pmv=adonis2(func_mat~Pressure, 
                        data=ind_df,
                        permutations =perm,
                        methods="bray", 
                        by = "margin")
                        
func_pressure_pmv
```
### Question 4: is there a significant difference in relative cover by functional group among sites within pressure levels?

```{r}
#SCRIPT WITH adonis()
set.seed(69)
perm=how(nperm=999)
func_site_pressure_pmv <- adonis(func_mat~Site, 
                           data=ind_df,
                           permutations =perm,
                           strata=ind_df$Pressure,
                           methods="bray",
                           by= "margin")

func_site_pressure_pmv$aov.tab

#ALTERNATIVE SCRIPT WITH adonis2()
set.seed(69)
h <- how(blocks = ind_df$Pressure, nperm = 999)
func_site_pressure_pmv <- adonis2(func_mat~Site,
                             data=ind_df,
                             permutations =h,
                             methods="bray",
                             by = "margin")

func_site_pressure_pmv
```

## Permutation test of multivariate homogeneity of groups dispersions (variances)

Theoretical explanation from Anderson, Marti, Ray N. Gorley, and Robert
K. Clarke. Permanova+ for primer: Guide to software and statisticl
methods. Primer-E Limited, 2008: *In the case of a one-way analysis, the
PERMANOVA test using permutations assumes only that the samples are
exchangeable under a true null hypothesis12. The assumption of
exchangeability is tantamount to assuming that the multivariate
observations (samples) are independent and identically distributed
(i.i.d.) under a true null hypothesis. Thus, although there are no
explicit assumptions regarding the distributions of the original
variables (they are certainly not assumed to be normally distributed),
independence and homogeneity of dispersions (in the space of the
resemblance measure) are directly implied by the permutation procedure.
Clearly, if samples have very different dispersions in different groups,
then they are not really exchangeable. Also, if the samples are
unequally correlated with one another (e.g., temporally or spatially),
then randomly shuffling them will destroy this kind of inherent
structure.*

```{r}
set.seed(69) 
densityplot(permustats(func_site_pmv)) #density plot of permutation results to to check if assumptions are met?

func_dist = vegdist(func_mat,method="bray")

func_bd=betadisper(func_dist,ind_df$Site)  #Implements Marti Anderson's PERMDISP2 procedure for the analysis of multivariate homogeneity of group dispersions (variances). betadisper is a multivariate analogue of Levene's test for homogeneity of variances 
str(func_bd)

anova(func_bd) #F-test on distribution of sites
```

### Permutest Implements a permutation-based test of multivariate homogeneity of group dispersions (variances)

```{r}
set.seed(69) 
permutest(func_bd) #permutative test on dispersion based on sites, if significance, PERMANOVA results may be driven by  dispersion issue
set.seed(69) 
permutest(func_bd, pairwise=TRUE) #pairwise of dispersion test
```

From Anderson, Marti, Ray N. Gorley, and Robert K. Clarke. Permanova+
for primer: Guide to software and statisticl methods. Primer-E Limited,
2008: \* If significant heterogeneity were detected by PERMDISP and
differences among groups were also detected using PERMANOVA, then the
latter could have been caused by differences in location, differences in
dispersion, or some combination of the two. \*

### Boxplot with within-site distance from centroids made with `ggplot()`

Create an input dataframe for the boxplot with mean relative cover by
category and depth within site

```{r}
func_bd$distances #Extract betadisper scores

distcent_df <- data.frame(Distance_to_centroid=func_bd$distances,
                          Site=func_bd$group) #create a df for the plot


distcent_df <- distcent_df %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

distcent_df[,2]<-ordered(distcent_df[,2],levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

distcent_df$Pressure<-ordered(distcent_df$Pressure,levels=c("Low", "Medium", "High"))

distcent_df_mean <- distcent_df %>%
  group_by(Site, Pressure) %>%
  summarise(Mean_dist = round(mean(Distance_to_centroid),2),
            SD = round(sd(Distance_to_centroid),2)) %>%
  ungroup()


colnames(distcent_df_mean) <- c("Site","Pressure","Mean distance to centroid","SD")
```

Boxplot with `geom_boxplot` from `ggplot2`

```{r}
box<- ggplot(data=distcent_df,aes(x=Site,y=Distance_to_centroid))+ 
  geom_boxplot(aes(fill=Pressure,colour=Pressure),alpha=0.3)+   
  geom_jitter(aes(color=Pressure),alpha=0.7)+
  theme_classic()+
  ylab(label="Distance to centroid")+
  theme(legend.position = "none")+ #remove legend
  scale_fill_manual(values=c("green", "yellow", "red"))+
  scale_color_manual(values=c("green", "yellow3", "red"))
box  
```

```{r, eval = F}
ggsave("func_boxplot.tiff", 
       plot = box,
       units="in", width=7, height=4, dpi=300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

### Post hoc test for PERMANOVA (`pairwise.adonis`)

```{r}
set.seed(69) 
func_pmv_phoc_holm= pairwise.adonis(func_mat,ind_df$Site,
                                  sim.function = "vegdist", 
                                  sim.method = "bray", 
                                  p.adjust.m = "holm", 
                                  reduce = NULL)

func_pmv_phoc_holm

set.seed(69) 
func_pmv_phoc_bonf= pairwise.adonis(func_mat,ind_df$Site,
                                  sim.function = "vegdist", 
                                  sim.method = "bray", 
                                  p.adjust.m = "bonferroni", 
                                  reduce = NULL)

func_pmv_phoc_bonf 
```

I can use Bonferroni because it is more conservative and it is the
adjustment adopted for the other analyses.

N.B. `pairwise.adonis` and `pairwise.adonis2` have different syntaxes!

# SIMPER: Similarity of percentage

From the vignette: *The simper functions performs pairwise comparisons
of groups of sampling units and finds the contribution of each species
to the average between-group Bray-Curtis dissimilarity. Although the
method is called simper, it really studied dissimilarities instead of
similarities (Clarke 1993).*

## SIMPER PER SITE:

```{r}
func_sim_site=simper(func_mat,group=ind_df$Site) #Distinguish categories which caused the differences in relative cover among sites
func_sim_site
summary(func_sim_site) #check which categories were the main driver of difference among sites
```

## SIMPER per depth

```{r}
func_sim_depth=simper(func_mat,group=ind_df$Depth) #Distinguish categories which caused the differences in relative cover among sites
func_sim_depth
summary(func_sim_depth, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

## SIMPER per pressure exposure
```{r}
func_sim_pres=simper(func_mat,group=ind_df$Pressure) #Distinguish categories which caused the differences in relative cover among sites
func_sim_pres
summary(func_sim_pres, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

# Session info

```{r}
sessionInfo()
```

# References

```{r}
#install.packages("report")
library(report)
report(sessionInfo())
```
