---
title: "Relative cover analysis"
author: "Claudia Campanini"
date: 
output: 
  html_document: 
    toc: yes
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load some basic libraries

```{r}
library(dplyr)
library(ggplot2)
library(readxl)   # to import excel datasets
library(writexl) # to export df to excel files (write_xlsx)
library(here)
```

# MEAN RELATIVE BENTHIC COVER BY CATEGORY

Load input file 

```{r}
cov_df <- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","plits_ext_df.xlsx"))
str(cov_df)
```
## PER SITE
Follow the link for explanation of dcast function https://seananderson.ca/2013/10/19/reshape/

```{r}
mean_rel_cov_site <- cov_df %>%
  group_by(Site, Category) %>%
  summarise(Mean_rel_cov=round(mean(Rel_cov),2))

library(reshape2) #needed for dcast function
mean_rel_cov_site <- dcast(mean_rel_cov_site, Site ~ Category)

mean_rel_cov_site <- mean_rel_cov_site %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High")
         )

mean_rel_cov_site$Site<-ordered(mean_rel_cov_site$Site,levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

mean_rel_cov_site <- mean_rel_cov_site[,c(1,11,2:10)] #re-ordering columns

mean_rel_cov_site$Pressure <- ordered(mean_rel_cov_site$Pressure,
                                      levels
                                      =c("Low","Medium","High"))

mean_rel_cov_site <- mean_rel_cov_site[order(mean_rel_cov_site$Site),]
mean_rel_cov_site
```
Export table with mean relative benthic cover by category per site within Depth level

```{r, eval= F}
write_xlsx(mean_rel_cov_site, here("Multi-taxa_data","PLITs","3-Clean_data","mean_rel_cov_site.xlsx"))
```

## PER SITE WITHIN Depth

```{r}
mean_rel_cov_df <- cov_df %>%
  group_by(Site_depth, Category) %>%
  summarise(Mean_rel_cov=round(mean(Rel_cov),2))

Site_depth <- mean_rel_cov_df$Site_depth

library(tidyr) # needed for separate_wider_delim function
mean_rel_cov_df <- mean_rel_cov_df %>%
  separate_wider_delim(Site_depth, "_", names = c("Site", "Depth")) %>%
  mutate(Site_depth=Site_depth)

rm(Site_depth)

mean_rel_cov_df <- 
  mean_rel_cov_df %>%
  mutate(Pressure = case_when(Site == "HI" | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low", 
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"
                              )
         ) 
  
mean_rel_cov_df$Site_depth <- as.factor(mean_rel_cov_df$Site_depth)

mean_rel_cov_df$Pressure <- ordered(mean_rel_cov_df$Pressure,
                                    levels=c("Low","Medium","High"))

mean_rel_cov_wide <- dcast(mean_rel_cov_df, 
                           Site_depth + Pressure ~ Category, 
                           value.var = "Mean_rel_cov")

mean_rel_cov_wide <- mean_rel_cov_wide[order(mean_rel_cov_wide$Site),]

mean_rel_cov_wide
```

Export table with mean relative benthic cover by category per site within Depth level

```{r, eval= F}
write_xlsx(mean_rel_cov_site, here("Multi-taxa_data","PLITs","3-Clean_data","mean_rel_cov_Site_depth.xlsx"))
```

## STACKED BARPLOTS

Load the color-blind-friendly palette with black:

```{r}
relcovPalette <- c("#009E73","#E69F00","#D55E00", "#56B4E9", "#F0E442", "#0072B2", "#CC79A7","#000000","#999999") #I changed the order compared to the original version
```

Stacked barplot of the mean relative benthic cover

N.B. I set a conditional labelling so that only mean relative cover
values \>0.15 are shown in the plot

```{r}
mean_rel_cov_df <- mean_rel_cov_df %>%                               # Replacing values
  mutate(Depth = replace(Depth, Depth == "Deep", "D")) %>%                     # Replacing values
  mutate(Depth = replace(Depth, Depth == "Shallow", "S"))

mean_rel_cov_df$Site<-ordered(mean_rel_cov_df$Site,levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

ggplot(mean_rel_cov_df, 
       aes(fill=Category, y=Mean_rel_cov, x=Depth, label=Mean_rel_cov)) + 
  geom_bar(position="fill", stat="identity",alpha=.8, color = "black")+
  facet_grid(~ Site)+
  theme_classic()+
  labs(x="Depth",
       y="Mean relative benthic cover",
       title= "Stacked bar chart of mean relative benthic cover \n by depth within site")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(labels = c("algae",
                                "boulder",
                                "hard coral",
                                "other",
                                "rubble",
                                "sand",
                                "soft coral",
                                "unknown",
                                "zoantharians"), values=relcovPalette)+
  scale_y_continuous(limits = c(0, 1), 
                     breaks = c(0, 0.25, 0.50, 0.75,1),
                     labels = c(0, 25, 50, 75,100))
```

```{r, eval= F}
ggsave("relcov_stacked.tiff", 
       units ="in", width = 5, height = 4, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

## STACKED BAR CHART BY PRESSURE

```{r}
mean_rel_cov_site <- cov_df %>%
  group_by(Site, Category) %>%
  summarise(Mean_rel_cov=round(mean(Rel_cov),2))

library(reshape2) #needed for dcast function
mean_rel_cov_site <- dcast(mean_rel_cov_site, Site ~ Category)

mean_rel_cov_site <- mean_rel_cov_site %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High")
         )

mean_rel_cov_site$Site<-ordered(mean_rel_cov_site$Site,levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

mean_rel_cov_site <- mean_rel_cov_site[,c(1,11,2:10)] #re-ordering columns

mean_rel_cov_site$Pressure <- ordered(mean_rel_cov_site$Pressure,
                                      levels
                                      =c("Low","Medium","High"))

mean_rel_cov_site <- mean_rel_cov_site[order(mean_rel_cov_site$Site),]
mean_rel_cov_site
```


```{r}
mean_rel_cov_pressure <- cov_df %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

mean_rel_cov_pressure <- mean_rel_cov_pressure %>%
  group_by(Pressure, depth, Category) %>%
  summarise(Mean_rel_cov=round(mean(Rel_cov),2))

mean_rel_cov_pressure$Pressure <- ordered(mean_rel_cov_pressure$Pressure,levels
                                      =c("Low","Medium","High"))
```

```{r}
ggplot(mean_rel_cov_pressure, 
       aes(x=depth, y=Mean_rel_cov,
           fill=Category, label=Mean_rel_cov)) + 
  geom_bar(position="fill", 
           stat="identity",
           alpha=.8, 
           color = "black")+
  facet_grid(~ Pressure)+
  theme_classic()+
  labs(x="Depth",
       y="Mean relative benthic cover (%)",
       fill="Benthic category")+
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")+
  guides(fill = guide_legend(title.position = "top"))+
  scale_fill_manual(labels = c("algae",
                                "boulder",
                                "hard coral",
                                "other",
                                "rubble",
                                "sand",
                                "soft coral",
                                "unknown",
                                "zoantharians"), 
                    values=relcovPalette)+
  scale_y_continuous(limits = c(0, 1), 
                     breaks = c(0, 0.25, 0.50, 0.75,1),
                     labels = c(0, 25, 50, 75,100)
                     )
```

```{r}
ggplot(mean_rel_cov_pressure, 
       aes(fill=Category, y=Mean_rel_cov, x=depth, 
           label=Mean_rel_cov)) + 
  geom_bar(position="fill", 
           stat="identity",
           alpha=.8, 
           color = "black")+
  facet_grid(~ Pressure)+
  theme_classic()+
  labs(x="Depth",
       y="Mean relative benthic cover (%)",
       fill="Benthic category")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(labels = c("algae",
                                "boulder",
                                "hard coral",
                                "other",
                                "rubble",
                                "sand",
                                "soft coral",
                                "unknown",
                                "zoantharians"), 
                    values=relcovPalette)+
  scale_y_continuous(limits = c(0, 1), 
                     breaks = c(0, 0.25, 0.50, 0.75,1),
                     labels = c(0, 25, 50, 75,100)
                     )
```

Export 

```{r, eval= F}
ggsave("relcov_stacked_pressure.tiff", 
       units ="in", width = 5, height = 6, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```


# nMDS

Load packages

```{r}
#VEGAN----
library(vegan)#load package
#install.packages("htmltools")
library(htmltools)
library(MASS)
library(dplyr)
library(readxl)   # to import excel datasets
```

## Format the data

Load relative benthic cover data

```{r}
#Relative benthic cover dataframe

cov_df <- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","Rel_cov_input.xlsx"))
str(cov_df)

cov_df <- cov_df %>% 
  rename(Depth = depth)

cov_df <- cov_df %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High"))

cov_df <- cov_df[ ,c(1:6, 25, 7:24)]

cov_df$Site<-ordered(cov_df$Site,levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

cov_df$Pressure<-ordered(cov_df$Pressure,levels=c("Low", "Medium", "High"))

cov_df[,c(1,3:6)] <- lapply(cov_df[,c(1,3:6)], as.factor)  ## as.factor() could also be used

cov_df[,8:25] <- lapply(cov_df[,8:25], as.numeric)

cov_df <- cov_df[,-c(17:25)]

str(cov_df)

#create a df with the independent variables
ind_df <- as.data.frame(cov_df[,c(1:7)])
str(ind_df)
```

Convert the database `cov_df`to a matrix

```{r}
cov_mat <- cov_df[,-c(1:7)]
rownames(cov_mat) <- cov_df$Transect
cov_mat <- as.matrix(cov_mat)
str(cov_mat)
class(cov_mat)
```
## Format and export data for Primer

```{r}
cov_mat_primer <- data.frame(t(cov_mat))
cov_mat_primer$Category <- colnames(cov_mat)
cov_mat_primer <- cov_mat_primer[,c(70,1:69)]
```

```{r}
ind_df_primer <- ind_df %>%
  mutate(Pressure_depth = paste(Pressure, Depth, sep='_')) 

ind_df_primer <- ind_df_primer[,-c(1,3,6)] # remove transect, transect depth, replicate
ind_df_primer <- ind_df_primer[,c(4,1:3,5)]
ind_df_primer <- data.frame(t(ind_df_primer))
ind_df_rownames <- rownames(ind_df_primer)
ind_df_primer$Factor <- ind_df_rownames
ind_df_primer <- ind_df_primer[,c(70,1:69)]
rm(ind_df_rownames)
```

```{r, eval =F}
#export each data frame to separate sheets in same Excel file
openxlsx::write.xlsx(list("cov_mat_primer" = cov_mat_primer,
                          "ind_df_primer" = ind_df_primer),
                     file = here("Multi-taxa_data","PLITs","3-Clean_data",
                   "cov_primer.xlsx")) 
```

## Run the Nonmetric Multidimensional scaling

```{r}
##metaMDS (isoMDS with random start, reduced  stress)----
set.seed(69) #random number to start
cov_nmds_br=metaMDS(cov_mat,distance = "bray",k=2) #Bray-Curtis resemblance matrix;k=2 means 2 dimensional
```

## Evaluate the nMDS mapping based on the stress and goodness of fit

```{r}
stressplot(cov_nmds_br) #evaluating NMDS mapping: stressplot
cov_nmds_br_gof=goodness(cov_nmds_br) #evaluating NMDS mapping: goodness of fit
cov_nmds_br_gof #smaller the value, the lower stress and better fit
plot(cov_nmds_br, type="t", main="goodness of fit") # create a plot for goodness of fit visualization:
points(cov_nmds_br, display="site", cex=cov_nmds_br_gof*100) #display the goodness of fit on the plot
#Stress lesser, better (>0.2 poor,0.2-0.1 moderate,0.1-0.05 good,<0.05 excellent); 
#K level higher, stress lesser
cov_nmds_br$stress
```
Stress = `r cov_nmds_br$stress` \< 0.1, not excellent (<0.05), but still a good ordination

Check Stress for each additional dimension

```{r}
k_vec=1:10
stress=numeric(length(k_vec))
relcov_dij=metaMDSdist(cov_mat, trace=FALSE)
set.seed(69)
for(i in seq_along(k_vec)) {
  cov_nmds_br_loop=metaMDSiter(relcov_dij, k=i, trace=FALSE)
  stress[i]=cov_nmds_br_loop$stress
}
plot(k_vec,stress,type="b",ylab="Stress",xlab="Dimensions") #select the K correspondent to the most drastic drop in stress
rm(k_vec)
```

## Plot the nMDS with basic R function `plot()`

```{r}
cov_nmds_br=metaMDS(cov_mat) #Run NMDS
plot(cov_nmds_br, type="text") #NMDS rough plot
cov_nmds_br$stress
```

## nMDS plot with `ggplot() and` with `geom_mark_ellipse()` from `ggforce`

Create input data frame with nMDS scores necessary for visualisation
with \`ggplot()\`

```{r}
cov_ndms_scores <- as.data.frame(scores(cov_nmds_br)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
cov_ndms_scores$transect <- rownames(cov_ndms_scores) # create a column of transect names
cov_ndms_scores$site <- ind_df$Site
cov_ndms_scores$Depth <- ind_df$Depth  #  add the grp variable created earlier
str(cov_ndms_scores)  #look at the data


NMDS1 <- cov_nmds_br$points[,1] ##also found using: > scores(cov_nmds_br)
NMDS2 <- cov_nmds_br$points[,2]
nmds_df<-cbind(ind_df, NMDS1, NMDS2)
rm(NMDS1,NMDS2)
```

nMDS plot

```{r}

str(cov_nmds_br)

library(ggplot2)
#install.packages("ggforce")
library(ggforce)

cov_ndms_ggplot <- ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Depth),size=3)+##separates overlapping points
  geom_mark_ellipse(aes(fill=Site, color=Site), expand = 0)+ # https://luisdva.github.io/rstats/Grouping-points/
  theme_classic()+
  ggtitle("nMDS of relative benthic cover")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  guides(color = guide_legend(order = 1,title="Site"),
         fill = guide_legend(order = 1,title="Site"),
         shape = guide_legend(order = 2,title="Depth"))+
  annotate("text", x=-0.45, y=-0.9, 
           label=paste('Stress =',round(cov_nmds_br$stress,3)),
           size=4)+
  annotate("text", x=-0.35, y=-1,
           label=bquote('Non-metric fit,'~ R^2~"= 0.998"),
           size=4) #add goodness of fit to plot

cov_ndms_ggplot
```

nMDS with ellipses based on the pressure level

```{r}
pal12 <- c("#000000","#004949","#009292","#ffb6db",
 "#490092","#006ddb","#b66dff","#b6dbff",
 "#920000","#924900","#24ff24","#ffff6d") #Cols from https://stackoverflow.com/questions/65013406/how-to-generate-30-distinct-colors-that-are-color-blind-friendly 

library(ggnewscale) #necessary for new_scale_color 

ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Depth),size=2.5)+##separates overlapping points
  scale_color_manual(values= pal12)+
  guides(color= guide_legend(order=1, title = "Site", ncol =2))+
  new_scale_color() +
  geom_mark_ellipse(aes(fill=Pressure, color=Pressure), expand = 0, alpha=0.2)+ # https://luisdva.github.io/rstats/Grouping-points/
  scale_color_manual(values= c("#009F81","#FFB331","#9F1111"))+
  scale_fill_manual(values= c("#009F81","#FFB331","#9F1111"))+
  theme_classic()+
  ggtitle("nMDS of relative benthic cover")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  scale_x_continuous(limits = c(-.8,1.8))+
  scale_y_continuous(limits = c(-.7,.9))+
  annotate("text", x=1.5, y=-0.6, 
           label=bquote(atop(Stress == .(round(cov_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.993")),
           size=3) #add goodness of fit to plot

rm(pal12)
```

```{r}
library(ggnewscale) #necessary for new_scale_color 

ggplot(nmds_df, aes(NMDS1, NMDS2))+
  geom_point(data=nmds_df, aes(NMDS1, NMDS2, color=Site, shape=Pressure),size=2.5)+##separates overlapping points
  guides(color= guide_legend(order=1, title = "Site", ncol =2))+
  new_scale_color() +
  geom_mark_ellipse(aes(fill=Pressure, color=Pressure), expand = 0, alpha=0.2)+ # https://luisdva.github.io/rstats/Grouping-points/
  scale_color_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  scale_fill_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  theme_classic()+
  ggtitle("nMDS of relative benthic cover")+
  theme(plot.title = element_text(size=14, face="bold",hjust=.5),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(size=10))+
  scale_x_continuous(limits = c(-.8,1.8))+
  scale_y_continuous(limits = c(-.7,.9))+
    annotate("text", x=1.5, y=-0.6, 
           label=bquote(atop(Stress == .(round(cov_nmds_br$stress,3)),
                             'Non-metric fit,'~ R^2~"= 0.993")),
           size=3) #add goodness of fit to plot
```

```{r, eval = F}
ggsave("ndms_relcov_pressure.tiff", 
       units ="in", width = 5, height = 4, dpi = 300, compression = 'lzw',
       path = here("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```

# BOXPLOTS FOR INGROUP VARIANCE

Load libraries

```{r}
#VEGAN----
library(vegan)#load package
library(MASS)
library(dplyr)

#install.packages("pairwiseAdonis")
library(devtools) #to install pairwiseAdonis from GitHub
library(pairwiseAdonis)
#install.packages("Rtools")
# library(remotes) #to use install_github
# install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)

library(readxl)   # to import excel datasets
```

Load and format data

```{r}
#Relative benthic cover dataframe

cov_df <- read_excel(here("Multi-taxa_data","PLITs","3-Clean_data","Rel_cov_input.xlsx"))
str(cov_df)

cov_df <- cov_df %>%
  mutate(Pressure = case_when(Site == 'HI' | 
                              Site == "S1" | 
                              Site == "S2" ~ "Low",
                              Site == "ZP" |
                              Site == 'ZA' | 
                              Site == "YO" | 
                              Site == "MI" ~ "Medium",
                              Site == 'SK' |
                              Site == 'SN' | 
                              Site == "AW" | 
                              Site == "G1" |
                              Site == "G2"~ "High")) %>%
  rename(Depth = depth)

cov_df <- cov_df[ ,c(1:6, 25, 7:24)]

cov_df$Site<-ordered(cov_df$Site,levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))

cov_df$Pressure<-ordered(cov_df$Pressure,levels=c("Low", "Medium", "High"))

cov_df[,c(1,3:6)] <- lapply(cov_df[,c(1,3:6)], as.factor)  ## as.factor() could also be used

cov_df[,8:25] <- lapply(cov_df[,8:25], as.numeric)

cov_df <- cov_df[,-c(17:25)]

str(cov_df)

#create a df with the independent variables
ind_df <- as.data.frame(cov_df[,c(1:7)])
str(ind_df)
```
Convert the database `cov_df`to a matrix

```{r}
cov_mat <- cov_df[,-c(1:7)]
rownames(cov_mat) <- cov_df$Transect
cov_mat <- as.matrix(cov_mat)
str(cov_mat)
class(cov_mat)
```

Create a df with explanatory (independent) variables

```{r}
ind_df <- as.data.frame(cov_df[,c(1:7)])
str(ind_df)
```

### Boxplot of within-site distance from centroid visualization with `ggplot()`

Create an input dataframe for the boxplot with mean relative cover by
category and depth within site

```{r}

relcov_bd$distances #Extract betadisper scores

distcent_df <- data.frame(Distance_to_centroid=relcov_bd$distances,Sites=ind_df$Site, Pressure=ind_df$Pressure) 
Sites <- relcov_bd$group
distcent_df[,2]<-factor(distcent_df[,2],levels=c("HI", "S1", "S2", "ZP", "ZA", "YO", "MI","SK", "SN", "AW", "G1", "G2"))
distcent_df[,3]<-factor(distcent_df[,3],levels=c("Low","Medium","High"))

distcent_mean_df <- distcent_df %>%
  group_by(Sites) %>% 
  summarise(`Mean distance to centroid`=round(mean(Distance_to_centroid),2),
         SD=round(mean(Distance_to_centroid),2)) %>%
  mutate(Pressure = case_when(Sites == 'HI' | 
                              Sites == "S1" | 
                              Sites == "S2" ~ "Low",
                              Sites == "ZP" |
                              Sites == 'ZA' | 
                              Sites == "YO" | 
                              Sites == "MI" ~ "Medium",
                              Sites == 'SK' |
                              Sites == 'SN' | 
                              Sites == "AW" | 
                              Sites == "G1" |
                              Sites == "G2"~ "High"))
  
distcent_mean_df <- distcent_mean_df[,c(1,4,2,3)]

relcov_site_box <- ggplot(data=distcent_df,aes(x=Sites,y=Distance_to_centroid))+ 
  geom_boxplot(aes(fill=Pressure,colour=Pressure),alpha=0.3)+
  geom_jitter(aes(fill=Pressure,color=Pressure),alpha=0.7)+
  scale_fill_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  scale_color_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  theme_classic()+
  labs(title="Boxplot of mean distance to centroid by site")+
  ylab(label="Distance to centroid")+
  theme(legend.position = "none", #remove legend
        plot.title  = element_text(hjust = 0.5, face= "bold", size =14),
        axis.text=element_text(size=14))

relcov_site_box

ggpubr:: ggarrange(relcov_site_box, 
                   gridExtra::tableGrob(distcent_mean_df), 
                   ncol = 2, widths = c(1,1))
```

```{r, eval=F}
ggsave("relcov_site_box.tiff", 
       units="in", width=12, height=4, dpi=300, compression = 'lzw',
       path = here ("Multi-taxa_data","PLITs","4-Visualization_outputs"))

ggsave("relcov_site_box.png", 
       units="in", width=12, height=4, dpi=300,
       path = here ("Multi-taxa_data","PLITs","4-Visualization_outputs"))
```


### Boxplot of within-site distance from centroid visualization with `ggplot()`

Create an input dataframe for the boxplot with mean relative cover by
category and depth within site

```{r}

relcov_bd_pres$distances #Extract betadisper scores

distcent_df_pres <- data.frame(
  Distance_to_centroid = relcov_bd_pres$distances, 
  Pressure = ind_df$Pressure) 

distcent_df_pres$Pressure  <- factor(distcent_df_pres$Pressure,
                                     levels = c("Low","Medium","High"))

distcent_mean_df_pres <- distcent_df %>%
  group_by(Pressure) %>% 
  summarise(`Mean distance to centroid`=round(mean(Distance_to_centroid),2),
         SD=round(mean(Distance_to_centroid),2))

relcov_pres_box <- ggplot(data=distcent_df_pres,aes(x=Pressure,y=Distance_to_centroid))+ 
  geom_boxplot(aes(fill=Pressure,colour=Pressure),alpha=0.3)+
  geom_point(aes(fill=Pressure,color=Pressure),alpha=0.7)+
  scale_fill_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  scale_color_manual(values= c("#009F81", "#FFB331","#9F1111"))+
  theme_classic()+
  labs(title="Boxplot of mean distance to centroid by pressure")+
  ylab(label="Distance to centroid")+
  theme(legend.position = "none", #remove legend
        plot.title  = element_text(hjust = 0.5, face= "bold", size =14),
        axis.text=element_text(size=14))

relcov_pres_box
```

# SIMPER

SIMPER: Similarity of percentage

From the vignette: *The simper functions performs pairwise comparisons
of groups of sampling units and finds the contribution of each species
to the average between-group Bray-Curtis dissimilarity. Although the
method is called simper, it really studied dissimilarities instead of
similarities (Clarke 1993).*

## SIMPER per site

```{r}

cov_mat_simper <- cov_mat[, colnames(cov_mat) != "unknown"] #remove "unknown" category from the matrix

relcov_sim=simper(cov_mat_simper, group=ind_df$Site) #Distinguish categories which caused the differences in relative cover among sites
relcov_sim
str(relcov_sim)
summary(relcov_sim) #check which categories were the main driver of difference among sites
```

## SIMPER per depth

```{r}
relcov_sim_depth=simper(cov_mat_simper,group=ind_df$Depth) #Distinguish categories which caused the differences in relative cover among sites
relcov_sim_depth
summary(relcov_sim_depth, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

## SIMPER per pressure exposure
```{r}
relcov_sim_pres=simper(cov_mat_simper,group=ind_df$Pressure) #Distinguish categories which caused the differences in relative cover among sites
relcov_sim_pres
summary(relcov_sim_pres, ordered=T, digits=3) #check which categories were the main driver of difference among sites
```

# Session info

```{r}
sessionInfo()
```

# References

```{r}
#install.packages("report")
library(report)
report(sessionInfo())
```

