---
title: "Extraction of relative cover by functional group from PLITs"
author: "Claudia Campanini"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

LOAD BASIC LIBRARIES

```{r}
library(readxl) #read xlsx files
library(writexl) #export to excel files
library(tidyverse) #aggregate fc et al.
library(tidyverse)
library(dplyr)
library(MASS)
```

# LOADING AND CLEANING OF RAW DATA

##IMPORT DATASHEETS FOR EACH SITE IN A SINGLE LIST OF LISTS

```{r}
files <- list.files(path = "C:/Users/claud/Documents/COMUNE/STUDIO.LAVORO/IMBRSea/Thesis/StatAnalysis/Multi-taxa_data/PLITs/1-Extraction/", pattern=".xlsx")
files
files <- files[-5]
class(files)
# Concat the directory to the file names
files <- str_c("C:/Users/claud/Documents/COMUNE/STUDIO.LAVORO/IMBRSea/Thesis/StatAnalysis/Multi-taxa_data/PLITs/1-Extraction/", files)
class(files)

#install.packages("rio")
library(rio)

plits_raw <- sapply (files, FUN = import_list,
              simplify = F, USE.NAMES = F)

names(plits_raw) = c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

rm(files)
```

## CLEAN PLITS OF EACH TRANSECT

```{r}
clean_plit_fun <- function(x) {
  x <- x[-c(1:2),c(1:5)]
  colnames(x) <-x[1,]
  x<- x[-1,]
  x$Point <-as.integer(x$Point)
  x$Cat_tot <-as.integer(x$Cat_tot)
  x$Func_tot <-as.integer(x$Func_tot)
  x$Category<- factor(x$Category, levels= c("algae","boulder","hard_coral","missing","other","rubble","sand",
                "shadow","soft_coral","sponge","unknown","zoanthids"))
  x$Functional_group<-as.factor(x$Functional_group)
  str(x)
  x
}

for (i in 1:12) {
  plits_raw[[i]] <- lapply (plits_raw[[i]],clean_plit_fun)
}

#alternative with map()
#for (i in 1:12) {plits_raw[[i]] <- map (plits_raw[[i]],clean_PLIT)}

rm(clean_plit_fun)
```
# HARD CORAL BY FUNCTIONAL GROUP 

## Subset raw data to hard coral only

```{r}
plits_hc <- vector("list", 12)
for (i in 1:12) {
    for (k in 1:length(plits_raw[[i]])) {
plits_hc[[i]][[k]] <- plits_raw[[i]][[k]] %>% 
  filter(Category == "hard_coral")
  }
}

names(plits_hc) = c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

depth_rep <- c("D1","D2","D3","S1","S2","S3")

for (i in 1:7) {
    for (k in 1:length(plits_hc[[i]])) {
    names(plits_hc[[i]]) = depth_rep
  }
}

for (i in 9:12) {
    for (k in 1:length(plits_hc[[i]])) {
    names(plits_hc[[i]]) = depth_rep
  }
}

names(plits_hc[[8]]) = c("S1","S2","S3")
```

Setting levels of `functional_group` removing functional groups with no records

```{r}
for (i in 1:12) {
    for (k in 1:length(plits_hc[[i]])) {
plits_hc[[i]][[k]]$Functional_group <- 
  factor(plits_hc[[i]][[k]]$Functional_group,                                        levels=as.character(unique(plits_hc[[i]][[k]]$Functional_group)[]))
  }
}
```

## Extraction

```{r}
ext_func_fun <- function(df, group_var,summary_var) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)
  
  df %>%
    filter(!is.na(!!group_var)) %>% 
    group_by(!! group_var) %>% 
    summarise(Abs_cov=sum(!! summary_var)) %>%
    mutate(Rel_cov = round(Abs_cov/sum(Abs_cov, na.rm = T)*100,1)) %>%
    complete(!! group_var)
    
}

hc_ext <- vector("list", 12)
for (i in 1:12) {
  hc_ext[[i]] <- lapply (plits_hc[[i]],ext_func_fun,
                             group_var = Functional_group, summary_var = Cat_tot)
}

names(hc_ext) = c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")
```

Add a column with the transect name to each transect (dataframe)

```{r}
#install.packages("wrapr") #https://search.r-project.org/CRAN/refmans/wrapr/html/add_name_column.html
library(wrapr)

for (i in 1:12) {
  hc_ext[[i]] <- add_name_column(hc_ext[[i]], "Depth_Rep")
}
```

Add a column with the site name to each transect (data frame)
```{r}
sites <- c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

for (i in 1:12) {
    for (k in 1:length(hc_ext[[i]])) {
    hc_ext[[i]][[k]]$Site <- as.factor(rep(sites[i],nrow(hc_ext[[i]][[k]])))
  }
}
```

Add a column for the depth level

```{r}
add_col_fun <- function(df, depth_rep) {
  depth_rep <- enquo(depth_rep)
  
  df %>% 
       mutate(Depth = case_when(!!depth_rep == 'D1' | 
                                  !!depth_rep == "D2" | 
                                  !!depth_rep == "D3"~ "Deep", 
                              !!depth_rep == 'S1' | 
                                  !!depth_rep == "S2" | 
                                  !!depth_rep == "S3"~ "Shallow"))
}

for (i in 1:12) {
  hc_ext[[i]] <- lapply (hc_ext[[i]],add_col_fun,
                             depth_rep= Depth_Rep)
}

```

## Merge data frames and formatting

```{r}
hc_ext_df <-list()
for (i in 1:12) {
  hc_ext_df[[i]] <- as.data.frame(do.call(rbind, hc_ext[[i]]))

}
hc_ext_df <- as.data.frame(do.call(rbind, hc_ext_df))

```

Add a column for Category

```{r}
hc_ext_df$Category <- as.factor(rep("hard_coral", nrow(hc_ext_df)))

hc_ext_df <- hc_ext_df[,c("Category","Functional_group", "Site", 
                            "Depth", "Depth_Rep", "Abs_cov", "Rel_cov")]

rm(hc_ext)
```

Hard coral cover by functional group within depth

```{r}
hc_per_depth <- hc_ext_df %>%
  group_by(Site, Depth, Category, Functional_group) %>%
  summarise(mean_cat_tot = round(mean(Abs_cov),0), 
            mean_perc=round(mean(Rel_cov),2), 
            sd=round(sd(Abs_cov),0))

```

# ALGAE COVER BY FUNCTIONAL GROUP

## Subset raw data to algae only 

```{r}
plits_alg <- vector("list", 12)
for (i in 1:12) {
    for (k in 1:length(plits_raw[[i]])) {
plits_alg[[i]][[k]] <- plits_raw[[i]][[k]] %>% 
  filter(Category == "algae")
  }
}

names(plits_alg) = c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

depth_rep <- c("D1","D2","D3","S1","S2","S3")

for (i in 1:7) {
    for (k in 1:length(plits_alg[[i]])) {
    names(plits_alg[[i]]) = depth_rep
  }
}

for (i in 9:12) {
    for (k in 1:length(plits_alg[[i]])) {
    names(plits_alg[[i]]) = depth_rep
  }
}

names(plits_alg[[8]]) = c("S1","S2","S3")

rm(depth_rep)
```

Setting levels of `functional_group` removing functional groups with no records

```{r}
for (i in 1:12) {
    for (k in 1:length(plits_alg[[i]])) {
plits_alg[[i]][[k]]$Functional_group <- 
  factor(plits_alg[[i]][[k]]$Functional_group,                                        levels=as.character(unique(plits_alg[[i]][[k]]$Functional_group)[]))
  }
}
```

## Extraction

```{r}
alg_ext <- vector("list", 12)
for (i in 1:12) {
  alg_ext[[i]] <- lapply (plits_alg[[i]],ext_func_fun,
                             group_var = Functional_group, summary_var = Cat_tot)
}

names(alg_ext) = c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

rm(ext_func_fun)
```

Add a column with the transect name to each transect (dataframe)

```{r}
#install.packages("wrapr") #https://search.r-project.org/CRAN/refmans/wrapr/html/add_name_column.html
library(wrapr)

for (i in 1:12) {
  alg_ext[[i]] <- add_name_column(alg_ext[[i]], "Depth_Rep")
}
```

Add a column with the site name to each transect (data frame)
```{r}
sites <- c("AW", "G1", "G2", "HI", "MI", "S1", "S2", "SK", "SN", "YO", "ZA", "ZP")

for (i in 1:12) {
    for (k in 1:length(alg_ext[[i]])) {
    alg_ext[[i]][[k]]$Site <- as.factor(rep(sites[i],nrow(alg_ext[[i]][[k]])))
  }
}

rm(sites)
```

Add a column for the depth level

```{r}
add_col_fun <- function(df, depth_rep) {
  depth_rep <- enquo(depth_rep)
  
  df %>% 
       mutate(Depth = case_when(!!depth_rep == 'D1' | 
                                  !!depth_rep == "D2" | 
                                  !!depth_rep == "D3"~ "Deep", 
                              !!depth_rep == 'S1' | 
                                  !!depth_rep == "S2" | 
                                  !!depth_rep == "S3"~ "Shallow"))
}

for (i in 1:12) {
  alg_ext[[i]] <- lapply (alg_ext[[i]],add_col_fun,
                             depth_rep= Depth_Rep)
}

rm(add_col_fun)
```

## Merge data frames and formatting

```{r}
alg_ext_df <-list()
for (i in 1:12) {
  alg_ext_df[[i]] <- as.data.frame(do.call(rbind, alg_ext[[i]]))

}
alg_ext_df <- as.data.frame(do.call(rbind, alg_ext_df))

rm(alg_ext, i , k)
```

Add a column for "Category"

```{r}
alg_ext_df$Category <- as.factor(rep("algae", nrow(alg_ext_df)))

alg_ext_df <- alg_ext_df[,c("Category","Functional_group", "Site", 
                            "Depth", "Depth_Rep", "Abs_cov", "Rel_cov")]
```

Algal cover by functional group within depth

```{r}
alg_per_depth <- alg_ext_df %>%
  group_by(Site, Depth, Category, Functional_group) %>%
  summarise(mean_cat_tot = round(mean(Abs_cov),0), 
            mean_perc=round(mean(Rel_cov),2), 
            sd=round(sd(Abs_cov),0))

```

# RELATIVE COVER BY FUNCTIONAL GROUPS 

Merging dataframes with hard-coral and algae cover per depth

```{r}
func_mean_depth_input <- rbind(hc_per_depth, alg_per_depth)
```

Merging dataframes with hard-coral and algae cover per transect

```{r}
func_per_transect <- rbind(hc_ext_df, alg_ext_df)
func_per_transect$Trans <- paste(func_per_transect$Site,
                                 func_per_transect$Depth_Rep, sep='_')
func_per_transect$Site_depth <- paste(func_per_transect$Site,
                                 func_per_transect$Depth, sep='_')
func_per_transect<- func_per_transect %>% 
       mutate(Rep = as.factor(case_when(Depth_Rep == 'D1' | Depth_Rep == "S1"~ 1, 
                              Depth_Rep == 'D2' | Depth_Rep == "S2"~ 2,
                              Depth_Rep == 'D3' | Depth_Rep == "S3"~ 3)))
```

## Rotate df

Follow the link for explanation of dcast function https://seananderson.ca/2013/10/19/reshape/

```{r}
library(reshape2)
func_input <- dcast(func_per_transect, 
                        Trans + Site + Depth + Rep + Site_depth  
                          ~ Functional_group, value.var = "Rel_cov")

func_input[is.na(func_input)] <- 0
```

# EXPORT INPUT FILES FOR RELATVE COVER BY FUNCTIONAL COVER ANALYSIS

```{r}
write_xlsx(func_input,"C:/Users/claud/Documents/COMUNE/STUDIO.LAVORO/IMBRSea/Thesis/StatAnalysis/Multi-taxa_data/PLITs/3-Relative_cover_functional_group/func_input.xlsx")

write_xlsx(func_mean_depth_input,"C:/Users/claud/Documents/COMUNE/STUDIO.LAVORO/IMBRSea/Thesis/StatAnalysis/Multi-taxa_data/PLITs/3-Relative_cover_functional_group/func_mean_depth_input.xlsx")
```

